---
title: "Forecast BTC price"
author: "Ignacio"
date: "10/03/2021"
output: html_document
---

```{r setup, include=FALSE}
## install packages

#install.packages("MASS")
#install.packages('forecast', dependencies = TRUE)
#install.packages("tidyverse")
#update.packages("MASS")
#install.packages("RColorBrewer")
#install.packages("gridExtra")
#install.packages("cowplot")
#install.packages("tidyverse")
#install.packages("GGally")
#library(Matrix)
##install.packages('latex2exp')  ## install LateX for plots
library(latex2exp)
library(gridExtra)
library(cowplot)
#library(viridis)
library(RColorBrewer)
library(tidyverse)
#library(forecast)
library(MASS)

```



```{r}
## Load dataset
d <- read.csv("Xday_merged.csv")
daux <- as.Date(d$Date)
d$Date <- daux
head(d)
```

```{r}
## quick plot

fig00 <- ggplot(data=d)+
  geom_line(aes(x=Date, y=BTC_Close))

fig00
```

```{r}
## Load Master Dataset
btc <- read.csv("BTC.csv")
btc$Date <- as.Date(btc$Date)
head(btc)
```


```{r}
## dimensions of the datasets
dim(d)
dim(btc)

```



```{r}
## define color palette for future plots
cpalette1 <- c("#CC6666", "#9999CC", "orange", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
## Build models
## log variables
lnClose <- log10(btc$Close)
lnS2F <- log10(btc$S2F)

df <- data.frame(lnClose, lnS2F, btc$SP500)

## polynomial models for S2F
m01 <- lm(data=df, formula=lnClose~I(lnS2F^1))
m02 <- lm(data=df, formula=lnClose~I(lnS2F^2))
m03 <- lm(data=df, formula=lnClose~I(lnS2F^3))
m04 <- lm(data=df, formula=lnClose~I(lnS2F^4))

summary(m01)
summary(m02)
summary(m03)
summary(m04)

```
```{r}
## 
names(m01)
dim(btc)
length(m01$residuals)
```


```{r}
## In this chunk we plot the curves for Bitcoin stock and supply
BTCs2f <- read.csv("BTCs2f.csv")
BTCs2f$Date <- as.Date(BTCs2f$Date)

## check scale
aux001 <- max(BTCs2f$BTC_stock); aux002 <- min(BTCs2f$BTC_stock)
aux003 <- max(log10(BTCs2f$BTC_flow)); aux004 <- min(log10(BTCs2f$BTC_flow))
a0 <- 2051048.39736512
b0 <- 13208261.98800764
aux_y <- a0*log10(BTCs2f$BTC_flow)+b0

## plot
#png("Fig-mined.png", height=600, width=1200)
fig05 <- ggplot(data=BTCs2f, aes(x=Date))+
  geom_line(aes(y=(BTC_stock), colour="BTC stock"), size=1)+
  #scale_y_continuous(sec.axis=sec_axis(~.*0.5e-4))+
  geom_line(aes(y=aux_y-0.5e7, colour="BTC mined"), size=1)+
  scale_colour_manual("", breaks=c("BTC stock", "BTC mined"), values=c(cpalette1[3], cpalette1[1]))+
  geom_vline(xintercept=as.Date("2021-02-01"), color="purple", linetype="dashed", size=0.5)

fig05
#fig05+theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()

```
```{r}
## print header for BTCS2F
head(BTCs2f)
```

```{r}
## In this chunk, we plot the results for different fitted models fotr the S2F model

## to make the plot easier, function to call the fit coefficients
beta_ <- function(lm){
  Bi <- lm$coefficients
  return(Bi)
}

## S2F


## plot
#png("Fig-S2F-poly  .png", height=600,width=1200)

fig03 <- ggplot(data=btc)+
  geom_line(aes(x=Date, y=Close, colour="Bitcoin"), size=1)+
  geom_line(aes(x=Date, y=beta_(m01)[1]+beta_(m01)[2]*S2F^1, colour="p=1"), size=1)+
  geom_line(aes(x=Date, y=beta_(m02)[1]+beta_(m02)[2]*S2F^2, colour="p=2"), size=1)+
  geom_line(aes(x=Date, y=beta_(m03)[1]+beta_(m03)[2]*S2F^3, colour="p=3"), size=1)+
  geom_line(aes(x=Date, y=beta_(m04)[1]+beta_(m04)[2]*S2F^4, colour="p=4"), size=1)+
  scale_y_continuous(trans="log10")+
  scale_colour_manual("", breaks=c("Bitcoin", "p=1","p=2","p=3", "p=4"), values=c("steelblue", cpalette1[1:4]))+
  labs(y="USD")+
  ggtitle(TeX(r"($x($S2F$)=\beta_0 + \beta_1 \cdot \log($S2F$)^p$)"))

fig03

#fig03 + theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()


## add the fitted S2F to the dataset
btc$logS2F_fit <- log10(beta_(m03)[1]+beta_(m03)[2]*(btc$S2F)^3)
names(btc)


```




```{r}
## R2 for the S2F fits
evalFit <- function(x, x2){
  RS_ <- na.remove(x - x2)    ## sum of residuals
  RSS_ <- sum(RS_^2)                      ## sum of squared residuals
  TSS_ <- sum((na.remove(x) - mean(na.remove(x)))^2)  ## total sum of squares
  RSE_ <- sqrt(RSS_/(length(x)-2))      ## residual standard error
  R2_ <- 1 - (RSS_/TSS_) ## R2 statistic
  MSE_ <- RSS_/length(na.remove(x))
  return(c(R2_, MSE_))
}

## dataframe for R2 values
evalm01 <- evalFit(btc$logClose, log10(beta_(m01)[1]+beta_(m01)[2]*btc$S2F^1))
evalm02 <- evalFit(btc$logClose, log10(beta_(m02)[1]+beta_(m02)[2]*btc$S2F^2))
evalm03 <- evalFit(btc$logClose, log10(beta_(m03)[1]+beta_(m03)[2]*btc$S2F^3))
evalm04 <- evalFit(btc$logClose, log10(beta_(m04)[1]+beta_(m04)[2]*btc$S2F^4))

try00 <- data.frame("logClose"=na.remove(btc$logClose), "p1"=evalm01[1], "p2"=evalm02[1], "p3"=evalm03[1], "p4"=evalm04)

## columns
model <- c("p1", "p2", "p3", "p4")
R2__ <- c(evalm01[1], evalm02[1], evalm03[1], evalm04[1])
MSE__ <- c(evalm01[2], evalm02[2], evalm03[2], evalm04[2])
table_R2 <- data.frame(model, R2__, MSE__)
table_R2 

```

```{r}
## columns
length((btc$Close))
```


```{r}
## columns
model <- c("p1", "p2", "p3", "p4")
R2__ <- c(evalFit(try00$logClose, try00$p1), evalFit(try00$logClose, try00$p2), evalFit(try00$logClose, try00$p3), evalFit(try00$logClose, try00$p4))
table_R2 <- data.frame(model, R2__)
table_R2 
```


```{r}
## Bitcoin and other markets
fig04 <- ggplot(data=btc)+
  geom_line(aes(x=Date, y=Close), color="steelblue")+
  geom_line(aes(x=Date, y=SP500), color="purple")+
  labs(y="USD")+
  scale_y_continuous(trans="log10")+
  geom_line(data=d, aes(x=Date, y=tesla_close), color="red")+
  geom_line(data=d, aes(x=Date, y=gold_close), color="black")+
  geom_line(data=d, aes(x=Date, y=lite_close), color="grey")

fig05 <- ggplot(data=btc)+
  geom_line(aes(x=Date, y=Close), color="steelblue")+
  scale_y_continuous(trans="log10")

fig04
fig05
```
```{r}
## header for other markets
head(btc)
```

# Time series for ARCH & GARCH
```{r}
## import libraries to perform ARCH & GARCH
#install.packages("FinTS")
#install.packages("rugarch")
#install.packages("tseries")
#install.packages("dynlm")
#install.packages("vars")
#install.packages("car")
#install.packages("forecast")

library(FinTS)          ## For ARCH models
library(rugarch)        ## For GARCH models
library(tseries)        ## For unit root test
library(dynlm)          ## For using lags in the model
library(vars)           ## For using var
#library(nlWaldTest)    ## For testing the non-linear Wald test
library(car)            ## For robust standard errors
library(forecast)
```


```{r}
#help(diff)

## convert to time series
btc$dClose <- c(NA, diff(btc$Close))
TS <- ts(btc[, c("Close","dClose")], start=c(2012, 01), frequency=360)

## plot
#png("Fig-ts.png", height = 600, width = 1200)
plot(TS)
#dev.off()

hist(TS[,1], freq=FALSE)
hist(TS[,2], freq=FALSE)

## Shapiro test
shapiro.test(TS)
```
We can see non-Gaussian distribution for the histograms. This is further verified using the Shapiro-Wilk test

We can see from the Shapiro normality test that the Bitcoin price is indeed not normally distributed. So the null hypothesis that the data follows a normal distribution is rejected and therefore we accept the alternative hypothesis that the time series follows a non-Gaussian distribution.


```{r}
## convert to log
btc$logClose <- log10(btc$Close)
btc$logdClose <- c(NA, diff(log10(btc$logClose)))

logTS <- na.remove(ts(btc[,c("logClose", "logdClose")], start=c(2012, 01), frequency=360))

## 
plot(logTS)

help(hist)

hist(logTS[,1], freq=FALSE, main="logClose")
hist(logTS[,2], freq=FALSE, main="logVelocity")

## Shapiro to the log prices
shapiro.test(logTS[,1])


```

```{r}
## Apply autocorrelation Function to the linear-scale data

## mean halving frequency
mhalv <- 1383 ## this is the average time difference in days between halvings

#png("Fig-ACF-pi.png", height=600, width=1200)
  
fig10 <- ggAcf(
  na.remove(btc$dClose),
  lag.max = 1200,
  type = c("correlation"),
  plot = TRUE,
  na.action = na.contiguous,
  demean = TRUE) +
  ggtitle("Autocorrelation-Function")+
  labs(x=TeX(r"(Lag ($\tau$))"))

fig10

#fig10 + theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()


## plot PI-cycle
#png("Fig-pi-cycle.png", height=600, width=1200)
fig02 <- ggplot(data=btc)+
  geom_ribbon(aes(x=Date, ymin=Short_Pi, ymax=Long_Pi, colour="Pi-cycle"), fill="grey85")+
  geom_line(aes(x=Date, y=Close, colour="Bitcoin"), size=1)+
  geom_line(aes(x=Date, y=Long_Pi), color="#00AFBB", size=1)+
  geom_line(aes(x=Date, y=Short_Pi), color="#FC4E07", size=1)+
  scale_y_continuous(trans="log10")+
  labs(y="USD")+
  scale_colour_manual("", breaks=c("Bitcoin", "Pi-cycle"), values=c("steelblue", "#CC79A7"))+
  ggtitle(TeX(r"($\pi$-cycle)"))
  

fig02
#fig02 + theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()

```


```{r}
## residulas and non-heteroscedasticity compared to S2F
re <- btc$logClose - btc$logS2F_fit
btc$re <- re
rss <- sum(re*re)

#png("Fig-residuals.png", height=600, width=1200)
fig08 <- ggplot(data=btc)+
  geom_point(aes(x=Date, y=re), color=cpalette1[6])+
  labs(y="Residuals")+
  geom_hline(yintercept=0, color="red")
fig08

#fig08 + theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()
```

This confirms a non-constant variance.

The estimate autoregression for a lag of 21 days is stationary then the 

```{r}
## Autocorrelation 

acf.log <- acf(logTS[,2], main="ACF", lag.max=200)

## P 
pacf.log <- pacf(logTS[,2])#, main="ACF", lag.max=200)

```

```{r}
## 
```



```{r}

## Autocorrelation ggplot
#png("Fig-ACF.png", height=600, width=1200)
  
fig07<- ggAcf(
  na.remove(btc$dClose),
  lag.max = 1200,
  type = c("correlation"),
  plot = TRUE,
  na.action = na.contiguous,
  demean = TRUE) +
  ggtitle("Autocorrelation-Function")+
  labs(x=TeX(r"(Lag ($\tau$))"))

fig07

#fig07 + theme(text=element_text(size=30), plot.title=element_text(hjust = 0.5)); dev.off()

```



```{r}
## Auto ARIMA
ar100 <- arima(btc$logdClose, c(1,0,0))
ar200 <- arima(btc$logdClose, c(2,0,0))

## ARMA
arma101 <- arima(btc$logdClose, c(1,0,1))

## print results
ar100
ar200
arma101

```
We can see that autocorrelation is significant intercept is not


```{r}
## now load GARCH model
help(ugarchfit)

## simpliest GARCH
g1s <- ugarchspec(variance.model=list(model="sGARCH", garchorder=c(1,1)), mean.model=list(armaorder=c(1,0)), distribution.model="std")
garch11 <- ugarchfit(g1s, data=logTS[,2])

garch11

```

```{r}
## convert to volatility
vole <- ts(garch11@fit$sigma^2, end=c(2021,02), frequency=360)


## plot volatility
png("Fig-garch11.png", height=600, width=1200)
plot(vole, xlab="Date", ylab=TeX(r"(volatility ($\sigma^2$))"), main="GARCH(1,1)",)
dev.off()
```


```{r}
## exponential GARCH
g1e <- ugarchspec(variance.model=list(model="eGARCH", garchorder=c(1,1)), mean.model=list(armaorder=c(1,0)), distribution.model="std",)
garch11e <- ugarchfit(g1e, data=logTS[,2])

garch11e
```

```{r}
## convert to volatility
vole2 <- ts(garch11e@fit$sigma^2, end=c(2021,02), frequency=360)


## plot volatility 
help(par)

#png("Fig-GARCH.png", height=1200, width=1300)
par(mfrow=c(3,1), cex.axis=2, cex.lab=3)
ts.plot(TS[,1], ylab="BTC")
ts.plot(TS[,2], ylab="returns")
ts.plot(vole, vole2, xlab="Date", ylab=TeX(r"(volatility ($\sigma^2$))"), col=c("red", "black"))
#dev.off()


```


